---
title: "MP_agents"
author: "EOL"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Mathcing Pennies Game

This markdown contains the code for the simulation two different strategies of the matching pennies game. The game is a simple two-player game where each player has to choose between two options (right or left). The winner is determined by whether the choices match or not.

## Strategy 1: Tired Win-Stay-Lose-Shift Agent
```{r}
# WSLS Agent function with noise (tiredness)
WSLSAgentNoise_f <- function(prevChoice, Feedback, noise){
  # Win-Stay-Lose-Shift Strategy
  if (Feedback == 1) {
    choice <- prevChoice
  } else if (Feedback == 0) {
    choice <- 1 - prevChoice
  }
  
  # Applying noise (tiredness)
  if (rbinom(1, 1, noise) == 1) {
    choice <- rbinom(1, 1, 0.5)
  }
  
  return(choice)
}
```

## Exponential Decay Agent
An agent that will decay its memory of previous choices over time.
It should also keep track of what choices are related to success using some weight parameter and make a choice in a given trial based on the highest weight. 

Moreover, it has a 10% chance of choosing randomly, reflecting tiredness. 

```{r alternative memory decay agent that uses weights}
MemoryDecayAgentWeight_f <- function(prevChoices, Feedback, decay, tiredness) {
  n <- length(prevChoices)

  if (rbinom(1, 1, tiredness) == 1) {
    choice <- rbinom(1, 1, 0.5) # if tired, choose randomly
  } else {
    decay_factor <- exp(-decay)
    prevChoices <- decay_factor * prevChoices

    # Calculate the exponentially decaying weights for each feedback
    weights <- exp(-decay * (1:n)) # small values for small values of n
    
    # Calculate the weighted proportion of positive feedbacks
    weighted_proportions <- cumsum(Feedback * weights) / cumsum(weights)
    
    # Choose the option with the highest weighted proportion
    choice <- ifelse(weighted_proportions[n] > 0.5, 1, 0)

    prevChoices <- c(prevChoices, choice)
  }

  # Return the updated choice and memory state
  return(list(choice = choice, memory = prevChoices))
}

```



## Tired WSLS and memory decay agent
```{r}
set.seed(1984)
trials <- 120

# Initialize vectors to store choices
self_tired_WSLS <- rep(NA, trials)
other_decay <- rep(NA, trials)
self_feedback <- rep(NA, trials)
other_feedback <- rep(NA, trials)

# RandomAgent_f was defined but not needed if initialized randomly

# Initialize the first choice randomly for both agents
self_tired_WSLS[1] <- sample(c(0, 1), 1)
other_decay_choices <- sample(c(0, 1), 1)
other_decay[1] <- other_decay_choices

# Simulate interactions between Tired Win-Stay-Lose-Shift Agent and Memory Decay Agent
for (i in 2:trials) {
  if (i == 2) {
    # First feedback is random for the first interaction
    Feedback_tired_WSLS <- sample(c(0, 1), 1)
    Feedback_decay <- 1 - Feedback_tired_WSLS
  } else {
    # Get feedback based on previous choices
    Feedback_tired_WSLS <- as.numeric(self_tired_WSLS[i - 1] == other_decay[i - 1])
    Feedback_decay <- 1 - Feedback_tired_WSLS
  }

  # Store feedback
  self_feedback[i - 1] <- Feedback_tired_WSLS
  other_feedback[i - 1] <- Feedback_decay
  
  # Make choices for the next trial
  self_tired_WSLS[i] <- WSLSAgentNoise_f(self_tired_WSLS[i - 1], Feedback_tired_WSLS, 0.2)
  
  # For the Memory Decay agent, we need to pass the cumulative vector rather than just the last feedback
  other_decay_decision <- MemoryDecayAgentWeight_f(other_decay_choices, other_feedback[1:(i-1)], 0.1, 0.1)
  other_decay_choices <- other_decay_decision$memory
  other_decay[i] <- other_decay_decision$choice
}

# Create a data frame for visualization
pacman::p_load(tibble, ggplot2)

# Create a data frame for visualization
df2 <- tibble(
  trial = seq(trials),
  self_tired_WSLS = self_tired_WSLS,
  other_decay = other_decay,
  Feedback_tired_WSLS = self_feedback
)

# Visualize the results
pl1 <- ggplot(df2, aes(trial)) +
  geom_line(aes(y = self_tired_WSLS, color = "Tired WSLS Agent"), size = 1) +
  geom_line(aes(y = other_decay, color = "Memory Decay Agent"), size = 1) +
  labs(title = "Tired WSLS Agent vs. Memory Decay Agent",
       y = "Choice", x = "Trial") +
  scale_color_manual(name = "Agents", values = c("Tired WSLS Agent" = "red", "Memory Decay Agent" = "blue")) +
  theme_classic()

# Calculate and plot the cumulative win rate for each agent
df2$cumulativeWinRateSelf <- cumsum(df2$Feedback_tired_WSLS) / seq_along(df2$Feedback_tired_WSLS)
df2$cumulativeWinRateOther <- (seq_along(df2$Feedback_tired_WSLS) - cumsum(df2$Feedback_tired_WSLS)) / seq_along(df2$Feedback_tired_WSLS)

pl2 <- ggplot(df2, aes(trial)) +
  geom_line(aes(y = cumulativeWinRateSelf, color = "Tired WSLS Agent"), size = 1) +
  geom_line(aes(y = cumulativeWinRateOther, color = "Memory Decay Agent"), size = 1) +
  labs(title = "Cumulative Win Rate of Tired WSLS Agent vs. Memory Decay Agent",
       y = "Cumulative Win Rate", x = "Trial") +
  scale_color_manual(name = "Agents", values = c("Tired WSLS Agent" = "red", "Memory Decay Agent" = "blue")) +
  theme_classic()

# Display the plots together
final_plt <- gridExtra::grid.arrange(pl1, pl2, ncol = 1)

# save the plot
ggsave("MP_agents.png", final_plt, width = 10, height = 10, units = "in")
```